\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-08-02 NAR_DRAFT.tex
changes

Assuming that the identity of the alternate allele and its allele frequency are independent of each other, and $\nu$ is uniformly distributed (i.e. $p(\nu) = 1$), we can rewrite equation \ref{eqn:1} as
\begin{equation}
  \label{eqn:2}
$$
  P(m,\nu) = p(m) \cdot \prod_{i=1}^{D} \textrm{f}_{m,\nu}(x_i).
$$
\end{equation}
Here we show how \batcave can be used to provide a tumor-specific estimate of the mutation probability $p(m)$.

@@

\begin{equation}
  \label{eqn:detailed_prior}
  $$
  \begin{array}{rcc}
  \mathrm{p}(\mathrm{S}) &=&  \mathrm{p}(m =j, M | C = i) \\
                            &=& \mathrm{p}(m = j \mid C = i) \mathrm{p}(C = i \mid M)\frac{p(M)}{p(C = i)},
  \end{array}
  $$
\end{equation}
where $\mathrm{p}(\mathrm{S})$ is the probability that a substitution of type $S$ has occurred.

@@

\subsection{Simulated data}
We generated six different tumor/normal pairs of two types, 3 each of 100X depth whole genomes and 500X whole exomes.
For each of the two types we simulated tumors with variants drawn from three different mutation profiles (Concentrated,Intermediate, and Diffuse) as described in Materials \& Methods and Figure \ref{NAR-sigfig}.
These 6 simulated tumor/normal pairs were used to assess the effect of sequencing depth and mutation profile type on the results obtained from \batcave.
\bkmcomment{Should there be a longer discussion of what sorts of tumors the profiles represent, or leave that for the methods only?}

@@

\begin{table*}[b]
\tableparts{%
\caption{This is a table with $\mu$, auroc, auprc, ICI, and maybe fraction called?}
\label{table:01}%
}{%
\begin{tabular*}{\textwidth}{@{}lllllllll@{}}
\toprule
Depth & Mutation profile & $\mu/\beta$ & AUROC & & AUPRC & & ICI & 
\\
& & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
\\
\colrule
100X whole genome & Concentrated & 3.6e-7 & .987 & .993 & .972 & .975 & -- & --
\\
100X whole genome & Intermediate & 3.2e-7 & .987 & .989 & .972 & .973 & -- & --
\\
100X whole genome & Diffuse & 3.2e-7 & .988 & .989 & .971 & .973 & -- & --
\\
500X whole exome & Concentrated & 3.6e-7 & .848 & .929 & .674 & .758 & .0585 & .0462
\\
500X whole exome & Intermediate & 3.6e-7 & .847 & .881 & .677 & .706 & .0587 & .0568 
\\
500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .0565 & .0576
\\
AML 31 & Actual & 3.6e-8 & .935 & .949 & .995 & .996 & 0.121 & 0.155
\\
\botrule
\end{tabular*}%
}
{$\mu$ = per-base mutation rate, AUROC/AUPRC = area under ROC/PRC curve, ICI = Integrated calibration index.}
\end{table*}

>>>>

Assuming that the identity of the alternate allele and its allele frequency are independent of each other, and $\nu$ is uniformly distributed (i.e. $p(\nu) = 1$), we can rewrite equation \ref{eqn:1} as
\begin{equation}
  \label{eqn:2}
$$
  P(m,\nu) = p(m) \cdot \prod_{i=1}^{\mathrm{d}} \textrm{f}_{m,\nu}(x_i).
$$
\end{equation}
Here we show how \batcave can be used to provide a tumor-specific estimate of the mutation probability $p(m)$.

@@

\begin{equation}
  \label{eqn:detailed_prior}
  $$
  \begin{array}{rcl}
  \mathrm{p}(\mathrm{S_{i,j}}) &=&  \mathrm{p}(m =j, M | C = i) \\
                            &=& \mathrm{p}(m = j \mid C = i) \mathrm{p}(C = i \mid M)\frac{p(M)}{p(C = i)},
  \end{array}
  $$
\end{equation}
where $\mathrm{p}(\mathrm{S_{i,j}})$ is the probability that a substitution of type $\mathrm{S_{i,j}}$ will occur.

@@

\subsection{Simulated data}
We generated six different tumor/normal pairs of two types, 3 each of 100X depth whole genomes and 500X whole exomes.
For each of the two types we simulated tumors with variants drawn from three different mutation profiles chosen to resemble melanoma (concentrated), a breast cancer (intermediate), and a lung cancer (diffuse)  as described in Materials \& Methods and Figure \ref{NAR-sigfig}.
These 6 simulated tumor/normal pairs were used to assess the effect of sequencing depth and mutation profile type on the results obtained from \batcave.

@@

\begin{table*}[b]
\tableparts{%
\caption{This is a table with $\mu$, auroc, auprc, ICI, and maybe fraction called?}
\label{table:01}%
}{%
\begin{tabular*}{\textwidth}{@{}lllllllll@{}}
\toprule
Depth & Mutation profile & $\mu$ & AUROC & & AUPRC & & ICI & 
\\
& & (estimated) & MuTect & \batcave & MuTect & \batcave & MuTect & \batcave
\\
\colrule
100X whole genome & Concentrated & 3.6e-7 & .987 & .993 & .972 & .975 & -- & --
\\
100X whole genome & Intermediate & 3.2e-7 & .987 & .989 & .972 & .973 & -- & --
\\
100X whole genome & Diffuse & 3.2e-7 & .988 & .989 & .971 & .973 & -- & --
\\
500X whole exome & Concentrated & 3.6e-7 & .848 & .929 & .674 & .758 & .0585 & .0462
\\
500X whole exome & Intermediate & 3.6e-7 & .847 & .881 & .677 & .706 & .0587 & .0568 
\\
500X whole exome & Diffuse & 3.6e-7 & .850 & .873 & .676 & .698 & .0565 & .0576
\\
Griffith et al. 2015 & Actual & 3.6e-8 & -- & -- & .995 & .996 & -- & --
\\
Shi et al. 2018 & Actual & 3.6e-8 & -- & -- & .972 & .972 & -- & --
\\
\botrule
\end{tabular*}%
}
{$\mu$ = per-base mutation rate, AUROC/AUPRC = area under ROC/PRC curve, ICI = Integrated calibration index.}
\end{table*}
